---
title: "Data Import and clean-up"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Project Intro

The project uses some data mining techniques with the Colorado Bridge Highway Inventory. The data has been taken from the Bridge Inspection Program of the Federal Highway Administration. For more details about the program see <https://www.fhwa.dot.gov/bridge/nbi.cfm>  

### Downloading data

26 years of inspection data is available (1992-2017). The code for dowloading data and extracting the one from Colorado is included below:

```{r eval = FALSE}
for(i in 1:26){ # 26 years availables
  year=(1991+i) # From 1992
  try(download.file(paste("https://www.fhwa.dot.gov/bridge/nbi/",year, 
                      "del.zip",sep=""),paste(
        "C:/Users/alexb/Google Drive/CVEN 6833 ADAT/zz Homeworks/Semester Project/NBI/",
        year,"del.zip",sep=""),method = "libcurl"))                 # download zip
  if(class(try(assign(paste("co",year,sep=""),read.delim2(unz(paste(
    "C:/Users/alexb/Google Drive/CVEN 6833 ADAT/zz Homeworks/Semester Project/NBI/",
    year,"del.zip",sep=""),                                        # extract Colorado
    grep("08fluna",unzip(paste(                                    # yy 1992-2000 +-
    "C:/Users/alexb/Google Drive/CVEN 6833 ADAT/zz Homeworks/Semester Project/NBI/",
    year,"del.zip",sep=""),list=TRUE)[,1],value = TRUE)),header=TRUE,sep=","))))[1]==
    "try-error"){
    try(assign(paste("co",year,sep=""),read.delim2(unz(paste(
      "C:/Users/alexb/Google Drive/CVEN 6833 ADAT/zz Homeworks/Semester Project/NBI/",
      year,"del.zip",sep=""),                                      # extract Colorado
      grep("CO",unzip(paste(                                       # yy 2000-2017 +-
        "C:/Users/alexb/Google Drive/CVEN 6833 ADAT/zz Homeworks/Semester Project/NBI/",
        year,"del.zip",sep=""),list=TRUE)[,1],value = TRUE)),header=TRUE,sep=",")))
  }else{}
}
```

### Setting up data

To help cleaning lat-long null data and storing in dec_deg format, we use a clean-up function

```{r eval = FALSE}
cleanup=function(dataframe){
dataframe$LAT_016<-ifelse(dataframe$LAT_016<=0,NA,dataframe$LAT_016) # NA to zero-neg values
dataframe$LONG_017<-ifelse(dataframe$LONG_017<=0,NA,dataframe$LONG_017) # NA to zero-neg values
dataframe=dataframe[complete.cases(dataframe$LAT_016)&complete.cases(dataframe$LONG_017),]
if(max(dataframe$LAT_016)>1e+6){
  for(i in 1:length(dataframe$LAT_016)){
    hour=as.double(substr(as.character(dataframe$LAT_016[i]),1,2))
    minute=as.double(substr(as.character(dataframe$LAT_016[i]),3,4))
    sec=as.double(substr(as.character(dataframe$LAT_016[i]),5,8))/100 # CHECK: Data w/ 2 digits only
    deg_min_sec=eval(paste(hour,minute,sec))
    deg_dec=as.double(conv_unit(deg_min_sec,from = 'deg_min_sec', to = 'dec_deg'))
    dataframe$LAT_016[i]=deg_dec
  }
  for(i in 1:length(dataframe$LONG_017)){
    hour=as.double(substr(as.character(dataframe$LONG_017[i]),1,3))
    minute=as.double(substr(as.character(dataframe$LONG_017[i]),4,5))
    sec=as.double(substr(as.character(dataframe$LONG_017[i]),6,9))/100 #CHECK: Data w/ 2 digits only
    deg_min_sec=eval(paste(hour,minute,sec))
    deg_dec=as.double(conv_unit(deg_min_sec,from = 'deg_min_sec', to = 'dec_deg'))
    dataframe$LONG_017[i]=-1*deg_dec
  }
}
  return(dataframe)
}
```

We can now apply this clean-up function to all available data (co1992 to co2017), and save it as a Rdata file.

```{r eval = FALSE} 
require(measurements) #eval = FALSE to save time knitting
for(i in 1:26){ # 26 years
  year=(1991+i) # From 1992
  assign(paste("co",year,sep=""),cleanup(eval(parse(text = paste("co",year,sep="")))))
  sprintf("year %s cleaned-up",year)
}
save(file="Colorado NBI.RData",list=grep("co",ls(),value = TRUE))
# next session / debugging RData can be re-loaded using
# load("Colorado NBI.RData")
```

### Plotting data

We'll plot some features to check the data, e.g. the sufficiency rating map for a given year. The bigger the square, the lower the rating is. 

```{r echo = FALSE}
# plot sufficiency ratio function
plot_suff=function(dataset){ #variable as character
  data=eval(parse(text=dataset))
  plot(data$LONG_017,data$LAT_016,xlim = c(-109,-102),ylim = c(37,41),
       pch=".",cex=1/as.numeric(levels(data$SUFFICIENCY_RATING))[data$SUFFICIENCY_RATING]*25)
  title(main = sprintf("Sufficiency rating year %s",substr(dataset,3,7)))
}
load("Colorado NBI.RData") #Has to be reloaded to be able to knit
par(mfrow=c(2,3))
for(i in 1:26){
  plot_suff(paste("co",1991+i,sep=""))
}

```

### Plotting with ggplot2
```{r}
require("tidyverse")
require("RColorBrewer")
ggplot_suff=function(dataset){ #rating fitted to parabola 0%=1, 50%=0.2, 100%=0
  df=eval(parse(text=dataset))[c("LONG_017","LAT_016","SUFFICIENCY_RATING","YEAR_BUILT_027")]
  df$SUFFICIENCY_RATING=as.numeric(levels(df$SUFFICIENCY_RATING))[df$SUFFICIENCY_RATING]
  ggplot(df,aes(df$LONG_017,df$LAT_016))+scale_size_continuous(range =c(0.01,4))+
  geom_point(aes(size=1.2e-4*df$SUFFICIENCY_RATING^2-2.2e-2*df$SUFFICIENCY_RATING+1,
  color=as.numeric(substr(dataset,3,7))-df$YEAR_BUILT_027))+scale_color_gradientn(
  colours =rev(brewer.pal(11,"RdYlBu"))) +coord_fixed(xlim = c(-109, -102), ylim = c(37, 41))+labs(title=sprintf("Sufficiency rating year %s",substr(dataset,3,7)),colour = "Age",size="S.R (recip)",x="Long",y="Lat")+theme(legend.position="bottom",legend.box = "horizontal") }
load("Colorado NBI.RData") #Has to be reloaded to be able to knit
gg=vector('numeric')
 for(i in 0:3){
     print(ggplot_suff(paste("co",1992+8*i,sep="")))
}
```
